---
title: "Homework 1"
author: "Ashley Kang"
date: "9/18/2018"
output: github_document
---

# Problem 1

```{r setup, include = FALSE}
library(tidyverse)
```

## Creating a dataframe
The chunk below creates a dataframe containing: 

* a sample of size 10 from a random uniform[0,5] distributed variable 
* a logical vector indicating whether elements of the sample are greater than 2 
* a (length-10) character vector 
* a (length-10) factor vector.

```{r create_prob1_df}
# Setting seed for reproducibility:
set.seed(1)

prob1_df = tibble(unif_samp = runif(n = 10, min = 0, max = 5), 
                  vec_logical = unif_samp > 2, 
                  vec_char = c("This", "is", "my", "first", "homework", 
                               "using", "R", "for", "data", "science"), 
                  vec_factor = factor(c("bad", "bad", "good", "mediocre", 
                                        "bad", "mediocre", "good", "bad", 
                                        "mediocre", "good"))
                  )


```

### Taking the mean of each variable

A numeric vector will return a mean. The mean of the random sample of size 10 from a uniform[0, 5] distribution is **`r mean(prob1_df$unif_samp)`**.
```{r mean_unif_samp, eval = FALSE}
samp_mean = mean(prob1_df$unif_samp)
```

A logical vector will return a mean, as values that are "TRUE" = 1 and values that are "FALSE" = 0. Averages can then be calculated for the translated numeric values. The mean for *vec_logical* is **`r mean(prob1_df$vec_logical)`**. 
```{r mean_vec_logical, eval = FALSE}
logical_mean = mean(prob1_df$vec_logical)
```

A character vector will not return a mean. Only vectors that are numeric or logical will return a mean. The mean for *vec_char* is **`r mean(prob1_df$vec_char)`**. 
```{r mean_vec_char, eval = FALSE}
char_mean = mean(prob1_df$vec_char)
```

A factor vector also will not return a mean. Only vectors that are numeric or logical will return a mean. The mean for *vec_factor* is **`r mean(prob1_df$vec_factor)`**. 
```{r mean_vec_factor, eval = FALSE}
factor_mean = mean(prob1_df$vec_factor)
```


### Explicitly converting variables

#### Converting into a numeric variable
The chunk below will apply as.numeric to the logical, character, and factor variables.

```{r converting_variable, eval = FALSE}
as.numeric(prob1_df$vec_logical)
as.numeric(prob1_df$vec_char)
as.numeric(prob1_df$vec_factor)
```

* The logical and factor variables were able to be converted into a numeric variable
    * As mentioned above, for a logical variable, "TRUE" = 1 and "FALSE" = 0, so it was converted as such
    * The factor variable was converted into a numeric variable, and the levels were translated as:
        * "bad" = 1
        * "good" = 2
        * "mediocre" = 3
* The character variable was not converted into a numeric one (warning message: NAs introduced by coercion). This is most likely because there is no inherent meaning/composition of the character variable (just composed of words which are composed of letters) so it cannot be converted numerically. Whereas factor variables are also composed of "words", there are levels associated, so it is able to be converted into a numeric variable.

#### Converting a character variable from character to factor to numeric; converting a factor variable from factor to character to numeric

```{r converting_char_and_factor, eval = FALSE}
# Convert character variable from character to factor to numeric
as.factor(prob1_df$vec_char)
as.numeric(as.factor(prob1_df$vec_char))

# Convert factor variable from factor to character to numeric
as.character(prob1_df$vec_factor)
as.numeric(as.character(prob1_df$vec_factor))
```

* The character variable was successfully converted to a factor variable, which created 10 levels, as each word within the character variable was unique. The factor variable was also converted to a numeric variable. The levels were sorted alphabetically and then assigned a numeric value (in ascending order) ranging from 1-10.

* The factor variable was successfully converted to a character variable. There are no longer "levels" associated with each word. The character variable was unable to be converted into a numeric variable, since there are no longer "levels" associated with each word, so NAs were introduced (by coercion).

