---
title: "Homework 1"
author: "Ashley Kang"
date: "9/18/2018"
output: html_document
---

# Problem 1

```{r setup, include = FALSE}
library(tidyverse)
```

## Creating a dataframe
The chunk below creates a dataframe containing a sample of size 10 from a 
random uniform[0,5] distributed variable, a logical vector indicating whether elements of the sample are greater than 2, a (length-10) character vector, and a (length-10) factor vector.

```{r create_data_frame}
set.seed(1)

prob1_df = tibble(unif_samp = runif(n = 10, min = 0, max = 5), 
                  vec_logical = unif_samp > 2, 
                  vec_char = c("This", "is", "my", "first", "homework", 
                               "using", "R", "for", "data", "science"), 
                  vec_factor = factor(c("less", "less", "greater", "greater", 
                                        "less", "greater", "greater", "greater", 
                                        "greater", "less"))
                  )


```

### Taking the mean of each variable
We will try to take the mean of each variable in the dataframe and see which variables return a mean value and which variables will not.

```{r mean, eval = FALSE}
mean(prob1_df$unif_samp)
mean(prob1_df$vec_logical)
mean(prob1_df$vec_char)
mean(prob1_df$vec_factor)
```

* Only numeric or logical variables will return a mean.
    + The mean for *unif_samp* is **`r mean(prob1_df$unif_samp)`**.
    + The mean for *vec_logical* is **`r mean(prob1_df$vec_logical)`**. 

* Character and factor variables will not return a mean.
    + The mean for *vec_char* is **`r mean(prob1_df$vec_char)`**.
    + The mean for *vec_factor* is **`r mean(prob1_df$vec_factor)`**.

### Explicitly converting variables

#### Converting into a numeric variable
The chunk below will explicitly convert logical, character, and factor 
variables into a numeric variable.

```{r converting_variable, eval = FALSE}
as.numeric(prob1_df$vec_logical)
as.numeric(prob1_df$vec_char)
as.numeric(prob1_df$vec_factor)
```
The logical and factor variables were able to be converted into a numeric variable,
but R was unable to turn a character variable into a numeric one (NAs introduced by
coercion).

#### Converting a character variable
The chunk below will convert the character variable from character to factor to numeric

```{r converting_character}
as.factor(prob1_df$vec_char)
as.numeric(prob1_df$vec_char)
```

R was able to convert the character variable to a factor variable and created 10 levels.
R was unable to convert the character variable to a numeric variable (NAs introduced by
coercion). 

#### Converting a factor variable
The chunk below will convert the factor variable from factor to character to numeric

```{r converting_factor}
as.character(prob1_df$vec_factor)
as.numeric(prob1_df$vec_factor)
```

R was able to convert the factor variable to both a character and numeric variable.
Now, there are 10 characters as opposed to 2 levels. There are only 2 numbers used - 
"1" to signify "greater" and "2" to signify "less".



# Problem 2

## Creating a dataframe for variables x and y

The chunk below creates a dataframe comprised of x (a random sample of size 1000 from a standard Normal distribution), y (a random sample of size 1000 from a standard Normal distribution), a logical vector indicating whether the x + y > 0, a numeric vector created by coercing the above logical vector, and a factor vector created by coercing the above logical vector.

```{r create_data_frame_2}
set.seed(1)

prob2_df = tibble(x = rnorm(1000), 
                  y = rnorm(1000), 
                  xy_logical = x + y > 0, 
                  num_coerc = as.numeric(xy_logical), 
                  fac_coerc = as.factor(xy_logical)
                  )
```
The data set is comprise of **`r nrow(prob2_df)` rows** and **`r ncol(prob2_df)` columns**.
The mean for *x* is **`r mean(prob2_df$x)`** and the median for *x* is **`r median(prob2_df$x)`**. 
The proportion of cases for which the logical vector is true is **`r mean(prob2_df$xy_logical)`**.

## Scatterplot of x and y

The chunk below will create a scatterplot of y vs x.

```{r scatter_yx}
ggplot(prob2_df, aes(x = x, y = y, color = xy_logical)) + geom_point()
ggplot(prob2_df, aes(x = x, y = y, color = num_coerc)) + geom_point()
ggplot(prob2_df, aes(x = x, y = y, color = fac_coerc)) + geom_point()
```

