---
title: "Homework 1"
author: "Ashley Kang"
date: "9/18/2018"
output: github_document
---

# Problem 1

```{r setup, include = FALSE}
library(tidyverse)
```

## Creating a dataframe
The chunk below creates a dataframe containing: 

* a sample of size 10 from a random uniform[0,5] distributed variable 
* a logical vector indicating whether elements of the sample are greater than 2 
* a (length-10) character vector 
* a (length-10) factor vector.

```{r create_prob1_df}
# Setting seed for reproducibility:
set.seed(1)

prob1_df = tibble(unif_samp = runif(n = 10, min = 0, max = 5), 
                  vec_logical = unif_samp > 2, 
                  vec_char = c("This", "is", "my", "first", "homework", 
                               "using", "R", "for", "data", "science"), 
                  vec_factor = factor(c("bad", "bad", "good", "mediocre", 
                                        "bad", "mediocre", "good", "bad", 
                                        "mediocre", "good"))
                  )


```

### Taking the mean of each variable

A numeric vector will return a mean. The mean of the random sample of size 10 from a uniform[0, 5] distribution is **`r mean(prob1_df$unif_samp)`**.
```{r mean_unif_samp, eval = FALSE}
samp_mean = mean(prob1_df$unif_samp)
```

A logical vector will return a mean, as values that are "TRUE" = 1 and values that are "FALSE" = 0. Averages can then be calculated for the translated numeric values. The mean for *vec_logical* is **`r mean(prob1_df$vec_logical)`**. 
```{r mean_vec_logical, eval = FALSE}
logical_mean = mean(prob1_df$vec_logical)
```

A character vector will not return a mean. Only vectors that are numeric or logical will return a mean. The mean for *vec_char* is **`r mean(prob1_df$vec_char)`**. 
```{r mean_vec_char, eval = FALSE}
char_mean = mean(prob1_df$vec_char)
```

A factor vector also will not return a mean. Only vectors that are numeric or logical will return a mean. The mean for *vec_factor* is **`r mean(prob1_df$vec_factor)`**. 
```{r mean_vec_factor, eval = FALSE}
factor_mean = mean(prob1_df$vec_factor)
```
