---
title: "Homework 1"
author: "Ashley Kang"
date: "9/18/2018"
output: html_document
---

# Problem 1

```{r setup, include = FALSE}
library(tidyverse)
```

## Creating a dataframe
The chunk below creates a dataframe containing a sample of size 10 from a 
random uniform[0,5] distributed variable, a logical vector indicating whether elements of the sample are greater than 2, a (length-10) character vector, and a (length-10) factor vector.

```{r create_data_frame}
set.seed(1)

unif_samp = runif(n = 10, min = 0, max = 5)
vec_logical = unif_samp > 2
vec_char = c("This", "is", "my", "first", "homework", 
               "using", "R", "for", "data", "science")
vec_factor = factor(c("less", "less", "greater", "greater", 
                        "less", "greater", "greater", "greater", 
                        "greater", "less"))

prob1_df = tibble(unif_samp, vec_logical, vec_char, vec_factor)


```

### Taking the mean of each variable
We will try to take the mean of each variable in the dataframe and see which variables return a mean value and which variables will not.

* Only numeric or logical variables will return a mean.
    + The mean for *unif_samp* is **`r mean(unif_samp)`**.
    + The mean for *vec_logical* is **`r mean(vec_logical)`**. 
* Character and factor variables will not return a mean.
    + The mean for *vec_char* is **`r mean(vec_char)`**.
    + The mean for *vec_factor* is **`r mean(vec_factor)`**.

### Explicitly converting variables

#### Converting into a numeric variable
The chunk below will explicitly convert logical, character, and factor 
variables into a numeric variable.

```{r converting_variable, eval = FALSE}
as.numeric(vec_logical)
as.numeric(vec_char)
as.numeric(vec_factor)
```
The logical and factor variables were able to be converted into a numeric variable,
but R was unable to turn a character variable into a numeric one (NAs introduced by
coercion).

#### Converting a character variable
The chunk below will convert the character variable from character to factor to numeric

```{r converting_character, eval = FALSE}
as.factor(vec_char)
as.numeric(vec_char)
```

R was able to convert the character variable to a factor variable and created 10 levels.
R was unable to convert the character variable to a numeric variable (NAs introduced by
coercion). 

#### Converting a factor variable
The chunk below will convert the factor variable from factor to character to numeric

```{r converting_factor, eval = FALSE}
as.character(vec_factor)
as.numeric(vec_factor)
```

R was able to convert the factor variable to both a character and numeric variable.
Now, there are 10 characters as opposed to 2 levels. There are only 2 numbers used - 
"1" to signify "greater" and "2" to signify "less".



# Problem 2

## Creating a dataframe for variables x and y

The chunk below creates a dataframe comprised of x (a random sample of size 1000 from a standard Normal distribution), y (a random sample of size 1000 from a standard Normal distribution), a logical vector indicating whether the x + y > 0, a numeric vector created by coercing the above logical vector, and a factor vector created by coercing the above logical vector.

```{r create_data_frame_2}
set.seed(1)

x = rnorm(1000)
y = rnorm(1000)
xy_logical = x + y > 0
num_coerc = as.numeric(xy_logical)
fac_coerc = as.factor(xy_logical)

prob2_df = tibble(x, y, xy_logical, num_coerc, fac_coerc)
```
The mean for *x* is **`r mean(x)`** and the median for *x* is **`r median(x)`**. 

## Scatterplot of x and y

The chunk below will create a scatterplot of y vs x.

```{r scatter_yx}
ggplot(prob2_df, aes(x = x, y = y, color = xy_logical)) + geom_point()
ggplot(prob2_df, aes(x = x, y = y, color = num_coerc)) + geom_point()
ggplot(prob2_df, aes(x = x, y = y, color = fac_coerc)) + geom_point()
```

